---
title: "rlevante_walkthrough"
output: html_document
---

---
title: "rlevante walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rlevante_walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Before you begin this user walkthrough, please see our [rlevante documentation](https://levante-framework.github.io/rlevante/index.html) for 1) how to install rlevante and 2) how to gain access to the data. 

This walkthrough uses an example dataset of 15 participants (levante-data-example) for teaching purposes. Access to the example dataset requires creating an account and joining the LEVANTE organization on Redivis, a data-sharing platform. For more information, see our [Researcher Site page on Data Access](https://researcher.levante-network.org/data).

## From Redivis to R

Permission to access the data is granted via [Redivis](https://redivis.com/), a data-sharing platform used for all LEVANTE datasets. Public releases of LEVANTE datasets on Redivis include at least four tables: participants, scores, surveys, and trials. 

**Table 1.** Redivis tables commonly read into data frames by rlevante

| Name         | Description                                                                                                                           | Additional information                                                                               |
|--------------|---------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------|
| participants | Information about child participants, where each row is a child participant.                                                          |  |
| scores       | Information about task ability scores across participants and tasks, where each row is a single score.                                | - [Overview of tasks](https://researcher.levante-network.org/measures/direct-child-measures)  <br> - [Guidelines for interpreting scores and details about scoring methods](https://researcher.levante-network.org/measures/scoring-and-psychometrics) |
| surveys      | Information about survey responses across participants and surveys, where each row is a single survey item response.                  | - [Child survey info](https://researcher.levante-network.org/measures/direct-child-measures/attitudes/thoughts-feelings) <br> -  [Caregiver survey info](https://researcher.levante-network.org/measures/caregiver-surveys) <br> - [Teacher survey info](https://researcher.levante-network.org/measures/teacher-survey)                                        |
| trials       | Information about trial level task responses across participants, tasks, and trials, where each row is a single trial-level response. | - [Overview of tasks](https://researcher.levante-network.org/measures/direct-child-measures)                                                                       |. 

## Using rlevante to load LEVANTE data
Each table from Redivis can be read into R as a data frame using a specific rlevante “get” function. Each “get” function takes at least two arguments, including a dataset name (required) and version number (defaults to the current version unless otherwise specified). 

First, load rlevante and other packages, as needed.
```{r setup}
library(rlevante)
library(dplyr)
library(ggplot2)
theme_set(theme_classic())
```

Then, use `get_participants` to obtain child participant ids, approximate birthdates, and any associated caregiver or teacher ids. 
```{r}
participants <- get_participants("levante-data-example:d0rt", version = "current")

participants |> count(dataset, sort = TRUE)
```

Use `get_scores` to obtain scored cognitive task data. 
```{r}
scores <- get_scores(data_source = "levante-data-example:d0rt")

scores |> count(dataset, sort = TRUE)

scores |> count(task_id, sort = TRUE)
```

```{r}
ggplot(scores, aes(x = age, y = score)) +
  facet_wrap(vars(task_id)) +
  geom_point()
```

Use `get_trials` to access trial-level data.

```{r}
trials <- get_trials("levante-data-example:d0rt")

trials |> count(dataset, sort = TRUE)

trials |> count(task_id, sort = TRUE)
```

Use `get_surveys` to access item-level survey data.
```{r}

surveys <- get_surveys("levante-data-example:d0rt")

surveys |> count(dataset, sort = TRUE)
```

Finally, researchers accessing their own LEVANTE data can use `get_raw_data` to access additional tables within private datasets.

```{r}

administrations <- get_raw_table(table_name = "administrations", data_source =  "levante-data-example-raw:bm7r")

administrations |> count(public_name, sort = TRUE)
```
