---
title: "rlevante_walkthrough"
output: html_document
---

---
title: "rlevante walkthrough"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rlevante_walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Before you begin this user walkthrough, please see our [rlevante documentation](https://levante-framework.github.io/rlevante/index.html) for 1) how to install rlevante and 2) how to gain access to the data. 

This walkthrough uses an example dataset of 15 participants (levante-data-example) for teaching purposes. Access to the example dataset requires creating an account and joining the LEVANTE organization on Redivis, a data-sharing platform. For more information, see our [Researcher Site page on Data Access](https://researcher.levante-network.org/data).

## From Redivis to R

Permission to access the data is granted via [Redivis](https://redivis.com/), a data-sharing platform used for all LEVANTE datasets. Public releases of LEVANTE datasets on Redivis include at least four tables: participants, scores, surveys, and trials. 

**Table 1.** Redivis tables commonly read into data frames by rlevante

| Name         | Description                                                                                                                           | Additional information                                                                               |
|--------------|---------------------------------------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------------------------|
| participants | Information about child participants, where each row is a child participant.                                                          |  |
| scores       | Information about task ability scores across participants and tasks, where each row is a single score.                                | - [Overview of tasks](https://researcher.levante-network.org/measures/direct-child-measures)  <br> - [Guidelines for interpreting scores and details about scoring methods](https://researcher.levante-network.org/measures/scoring-and-psychometrics) |
| surveys      | Information about survey responses across participants and surveys, where each row is a single survey item response.                  | - [Child survey info](https://researcher.levante-network.org/measures/direct-child-measures/attitudes/thoughts-feelings) <br> -  [Caregiver survey info](https://researcher.levante-network.org/measures/caregiver-surveys) <br> - [Teacher survey info](https://researcher.levante-network.org/measures/teacher-survey)                                        |
| trials       | Information about trial level task responses across participants, tasks, and trials, where each row is a single trial-level response. | - [Overview of tasks](https://researcher.levante-network.org/measures/direct-child-measures)                                                                       |. 

## Using rlevante to load LEVANTE data
Each table from Redivis can be read into R as a data frame using a specific rlevante “get” function. Each “get” function takes at least two arguments, including a dataset name (required) and version number (defaults to the current version unless otherwise specified). 

First, load rlevante and other packages, as needed.
```{r setup}
library(rlevante)
library(dplyr)
library(ggplot2)
theme_set(theme_classic())
```

Then, use `get_participants` to obtain child participant ids, approximate birthdates, and any associated caregiver or teacher ids. 

`get_participants` is the first rlevante function in this code. When you run it, you'll be prompted to authenticate with a pop-up browser on Redivis (an academic database manager that we use at LEVANTE) to ensure that you have the right permissions to access the dataset. Subsequent rlevante calls to this dataset within the same session should not require additional authentication.

```{r}
# Pull participant data from Redivis and save it to a data frame called "participants"
participants <- get_participants(data_source = "levante-data-example:d0rt", version = "current") 

# Now let's do some preliminary checks on the participants data
participants |> count(site, sort = TRUE) # returns the number of participants from each site — the levante-data-example and private site data releases only contain data from a single site, but the public data releases contain many sites/datasets stapled together
```

Use `get_scores` to obtain scored cognitive task data. 
```{r}
# Pull scored data from Redivis and save it to a data frame called "scores"
scores <- get_scores(data_source = "levante-data-example:d0rt", version = "current") 

# Now let's do some preliminary checks on the scores data
scores |> count(task_id, sort = TRUE) # returns the number of scores for each task. for a list of item_task abbreviations, see Table 1 here: https://researcher.levante-network.org/measures/direct-child-measures
```

```{r}
# This plot returns scores by age fro each task.
ggplot(scores, aes(x = age, y = score)) +
  facet_wrap(vars(task_id)) +
  geom_point()
```

Use `get_trials` to access trial-level data.

```{r}
# Pull trial-level data from Redivis and save it to a data frame called "trials"
trials <- get_trials(data_source = "levante-data-example:d0rt", version = "current") 

# Now let's do some preliminary checks on the trials data
trials |> count(site, sort = TRUE) # returns the number of trials per site

trials |> count(task_id, sort = TRUE) # returns the number of trials per task
```

Use `get_surveys` to access item-level survey data.
```{r}
# Pull survey data from Redivis and save it to a data frame called "surveys"
surveys <- get_surveys(data_source = "levante-data-example:d0rt", version = "current") 

# Now let's do some preliminary checks on the trials data
surveys |> count(site, sort = TRUE) # returns the number of surveys per site
```

Use `get_item_parameters` to access the IRT item parameters used in LEVANTE scoring.
```{r}
# Pull up-to-date item parameters from Redivis
item_parameters <- get_item_parameters()

item_parameters |> count(item_task, sort = TRUE) # returns the number of parameters per task
```

Finally, researchers accessing their own LEVANTE data can use `get_raw_data` to access additional tables within private datasets. `get_raw_data` is a general function that can pull any table that you specify, provided that the table exists. When pulling raw data, please note that some of the table names are different than the processed data. Check your dataset page to ensure that the table exists. 

```{r}
# Pull data from any table within from Redivis – in this case, we're pulling the administrations table (each assignment is an administration)
administrations <- get_raw_table(table_name = "administrations", data_source =  "levante-data-example-raw:bm7r") 

administrations |> count(public_name, sort = TRUE) # returns the number of completions per assignment, based on that assignment's name
```
