---
title: "Scoring and the model registry"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scoring and the model registry}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # the code below downloads models and data from Redivis (and fits IRT
  # models), so it is shown for reference but not run when the vignette builds
  eval = FALSE
)
```

The [walkthrough vignette](rlevante_walkthrough.html) covers downloading
*already-scored* data with `get_scores()`. This vignette is for the next step:
understanding **how** those scores are produced, and how to reproduce or audit
them yourself using the LEVANTE **model registry**.

```{r setup}
library(rlevante)
library(dplyr)
```

## How LEVANTE scores are produced

Most cognitive-task scores in LEVANTE are IRT ability estimates (θ). They are
produced by fitting item-response models once (calibration) and then scoring
each child's trial data against the appropriate fitted model. The fitted models
and the bookkeeping needed to find them live in the
`levante_metadata_scoring` dataset on Redivis, which rlevante reads through a
handful of functions:

| Function | Returns |
|---|---|
| `fetch_scoring_table()`   | which model specification applies to each (task, dataset) |
| `fetch_registry_table()`  | an index mapping model files to Redivis file ids |
| `fetch_registry_dir()`    | a Redivis directory handle for downloading model files |
| `score()`                 | scores one (task, dataset) using the above |

End to end, scoring a task flows like this:

```
get_trials(data_source)             # raw trial data
   |> recode_trials()               # make trials scoring-ready (see below)

fetch_scoring_table()               # look up the model spec for (task, dataset)
fetch_registry_dir()                # handle for downloading the fitted model

score(task, dataset, trials, runs,  # -> downloads the ModelRecord, shapes the
      scoring_table, registry_dir)  #    data, rebuilds the model, scores it
```

## Pinning a dataset version

Every rlevante "get" function takes a `data_source` and a `version`. For
reproducible analyses you should pin a **fully qualified reference** of the form
`name:hash:version` rather than relying on `version = "current"`, which silently
tracks the latest release.

```{r}
# tracks whatever the latest release happens to be
scores_current <- get_scores("levante_data_example:d0rt", version = "current")

# pinned: name + persistent dataset id (hash) + dataset version
scores_pinned <- get_scores("levante_data_example:d0rt:v1_0")
```

The persistent `hash` (e.g. `d0rt`) is stable even if the dataset is renamed;
the trailing `:v1_0` pins the version. You can find both on the dataset's
Redivis page under **API Information** → **Qualified references** (see the
["Reference identifiers" section of the README](https://levante-framework.github.io/rlevante/index.html)).
When you call a "get" function with `version = "current"`, rlevante emits a
message reporting the qualified reference it resolved to, which you can copy to
pin future runs.

## Trials must be recoded before scoring

`get_trials()` returns a `correct` column, but that column is **not yet
scoring-ready**. The calibrated models were fit on data passed through
`recode_trials()` first, so you must apply it before scoring or you will get
subtly wrong numbers.

```{r}
trials <- get_trials("levante_data_example:d0rt") |>
  recode_trials()
```

`recode_trials()` performs several task-specific corrections, including:

- **Slider items** (e.g. number line): marks a response correct when it falls
  within `slider_threshold` (default 0.15) of the target, and sets the chance
  level to `1 / slider_threshold / 100` accordingly.
- **Hearts & Flowers (`hf`)**: codes too-fast (<200 ms) and too-slow (>2000 ms)
  responses as incorrect and labels each trial start/stay/switch.
- **Same/Different Selection (`sds`)** and **Theory of Mind (`tom`)**: recodes
  correctness and disaggregates items.
- **Known item-key fixes**: e.g. `math_subtract_37_24` is rescored against the
  corrected answer.

## Rescoring a task yourself

Putting it together, here is how to reproduce the scores for one task and
dataset and compare them against the published values. The exported `score()`
function is the high-level entry point — it looks up the model spec, downloads
the fitted `ModelRecord`, shapes the data, and scores it.

```{r}
scoring_table <- fetch_scoring_table()
registry_dir  <- fetch_registry_dir()

task <- "math"
dataset <- "pilot_uniandes_co_bogota"

# trial (and, for adaptive tasks, run) data for this task + dataset
trials_task <- get_trials("levante_data_example:d0rt") |>
  recode_trials() |>
  filter(item_task == task, dataset == !!dataset)

my_scores <- score(
  task = task, dataset = dataset,
  trials = trials_task, runs = NULL,
  scoring_table = scoring_table, registry_dir = registry_dir
)

# compare to the published scores
published <- get_scores("levante_data_example:d0rt") |>
  filter(task_id == task, dataset == !!dataset)

my_scores |>
  inner_join(published, by = "run_id", suffix = c("_mine", "_pub")) |>
  summarise(cor = cor(score_mine, score_pub))
```

## Working with a `ModelRecord`

`score()` downloads a fitted model as a `ModelRecord` object. If you want to
inspect a model directly — for example to look up item parameters, or to
validate a rescore — you can download one yourself from the registry. The
accessors and slots you will need:

```{r}
# look up and download the ModelRecord for a (task, dataset)
spec <- fetch_scoring_table() |>
  filter(item_task == "math", dataset == "pilot_uniandes_co_bogota") |>
  as.list()
mod_rec <- get_registry_file(model_spec_filename(spec), fetch_registry_dir())
```

| Accessor | Returns |
|---|---|
| `items(mod_rec)`        | the model's item names, **in the order the model uses them** |
| `model_class(mod_rec)`  | `"SingleGroupClass"` or `"MultipleGroupClass"` |
| `model_vals(mod_rec)`   | the `mirt` parameter table for the fitted model |
| `scores(mod_rec)`       | the **calibration-time** EAP scores (`run_id`, `ability`, `se`) |

Some details are exposed as S4 slots:

| Slot | Contents |
|---|---|
| `mod_rec@scores`      | calibration EAP scores (same as `scores(mod_rec)`) |
| `mod_rec@data`        | the response matrix used to fit the model |
| `mod_rec@groups`      | the group label of each calibration row |
| `mod_rec@group_names` | the model's group names (for multigroup models) |

Two invariants are worth knowing. First, `mod_rec@data`'s columns are
guaranteed to be in `items(mod_rec)` order — scoring code must align new data to
this order before calling `mirt::fscores()`, which matches columns by position.
Second, `scores(mod_rec)` holds the EAP scores computed at calibration time on
the full data matrix, so it is a ground-truth reference you can validate a
rescore against:

```{r}
# rescoring the calibration runs should reproduce scores(mod_rec)$ability
scores(mod_rec)
```
